Disclosing funding sources for open access publication fees: the Open APC

### 2016-05-25 article.Rmd
research question paragraph and methods completed

The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015. Drawing on self-reported data that German institutions openly shared within the Open APC initiative, we focused our analysis on the amount of institutional publication fee spending, and compared these findings with that of related efforts in Austria and the UK. We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.

@@

We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations. In addition to administrative data about the amount paid per article including value added tax and the year of invoicing, the Open APC initiative has begun to share information about whether an article was published in a fully or hybrid open access journal. Bibliographic metadata recorded in the data-set consisted of the Digital Object Identifier (DOI) to uniquely identify an article, and information about publisher and journal titles that were automatically retrieved from CrossRef when available.

Data collection in this study involved obtaining cost data from related open data initiatives and retrieving bibliographic metadata updates. To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets provided by the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc. All data-sets were openly available through a figshare collection. We obtained the size  of funding for all publications and fully open access journal articles. The Euro foreign exchange reference rates from the European Central bank were consulted to convert costs reported in GBP to Euro.

To provide a current view on open access publishing, DOIs reported in the Open APC data-set served us as a basis to re-collect bibliographic metadata from CrossRef on May 19, 2016. Since the beginning of data normalisation using CrossRef by the Open APC initiative in 2014, CrossRef has provided incremental metadata updates in which ongoing mergers of publishing houses or name changes are reflected. A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing announced on May 6, 2015, that now operate under Springer Nature. In contrast to this, the Open APC initiative gathered publisher and journal titles shortly after an institution provided its expenditures.

As a client to access the CrossRef web service,  we used the R package rcrossref [@rcrossref] developed and maintained by the rOpenSci initiative.[^6] With the function `cr_cn`, we requested the linking resource `application/vnd.crossref.unixsd+xml` used  to support text mining activities.[^5] At the time of our study, this XML format distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals. It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014]. In cases where no bibliographic information could be obtained, we used the Open APC values. Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.

- statistical calculations

>>>>

The aim of this study was to examine how much German universities and research organisations spent on open access publication fees until 2015. Drawing on self-reported data that German institutions openly shared through the Open APC initiative, the analysis focused on the amount of institutional payments for publication fees, and compared these findings with those from related Austrian and British initiatives. We also asked how thoroughly self-reported articles were indexed in CrossRef, a DOI minting agency for scholarly literature, and analysed how institutional spending per articles was distributed over publishers and journal titles.

@@

We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations. In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI.

Data collection in this study involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc. For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles. In the case of FWF, we gathered the cost information from accompanying reports. We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.

Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's. Although the Open APC initative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information. 
However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes. A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that now operates as Springer Nature. To reflect these changes in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.

As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API. We requested the XML-based format `application/vnd.crossref.unixsd+xml`, which distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals. It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014]. In cases where no bibliographic information could be obtained, we used the Open APC values. Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.
